---
title: "Is there a surge of bankruptcies due to Covid-19?"
author: "Jorgen Reinnes"
date: "10/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Because of the Covid pandemic there has been a difficult time for some groups of companies as they have lost business. How does it compare to 2019?

```{r fetchData, echo=FALSE}
url <- "https://w2.brreg.no/kunngjoring/kombisok.jsp?id_region=0&id_niva1=51&id_niva2=56&id_bransje1=0"

library('xml2')
library('rvest')

scrape_search <- function(url, from_date, to_date) {
  data <- data.frame(county=rep(NA, 5000), company=rep(NA, 5000), orgNumber=rep(NA, 5000), date=rep(NA, 5000), type=rep(NA, 5000))
  found <- 0
  
  # Get website html
  webpage <- read_html(paste(url, "&datoFra=", from_date, "&datoTil=", to_date, sep = ""))
  
  # Get the data of search
  rows <- webpage %>% html_nodes("table:nth-of-type(4) table tr")
  
  # Loop through rows and clean data length(rows)
  county = "TBD"
  # Skip first 5 rows as they contain no data
  for (i in 6:length(rows)){
    tds <- rows[i] %>% html_nodes("td")
    if (length(tds) == 2){
      # County name Remove spaces and new lines
      county <- gsub("^\\s+|\\s+$", "", rows[i] %>% html_text())
    } else if (length(tds) > 3) {
      # Brokerage
      orgNumber <- trimws(gsub("\n", "", tds[4] %>% html_text()))
      # Only record companies which will have 4th td string length of atleast 9 chars:
      if (nchar(orgNumber) == 11){
        found <- found + 1
        data$county[found] <- county
        data$company[found] <- trimws(gsub("\n", "", tds[2] %>% html_text()))
        data$orgNumber[found] <- orgNumber
        data$date[found] <- tds[6] %>% html_text()
        data$type[found] <- trimws(gsub("\n", "", tds[8] %>% html_text()))
      }
    } else {
      # Ignore white space rows
    }
  }
  return(data[1:found,])
}

result <- data.frame()
# For each month search given the 5000 results limit until today
library(lubridate)
from_date <- ymd("2019-01-01")
now_date <- today()
while (from_date <= now_date){
  to_date <- from_date %m+% months(1) %m+% days(-1)
  result <- rbind(result, scrape_search(url, from_date %>% format("%d.%m.%Y"), to_date %>% format("%d.%m.%Y")))
  from_date <- from_date %m+% months(1)
}

# Cleanup data
library("dplyr")

# Fix date type
result <- result %>% mutate(date = as.Date(date, format="%d.%m.%Y")) %>%
  # Filter out data from outside the country
  filter(county!="Utenlands") %>% filter(type=="Konkursåpning")

# Add year, and format date day to 01 for all
preAggregatedData <- result %>% mutate(year=format(date, "%Y")) %>% mutate(month=format(date, "%m"))

preCovidYear <- preAggregatedData %>% filter(year == "2019")
covidYear <- preAggregatedData %>% filter(year == "2020")


```

## 2019 vs 2020 Bankruptcies Per County

```{r plots, echo=FALSE}
require(data.table)
require(ggplot2)
require(tidyverse)

fill_empty_months <- function(df, months) {
  # df has all 12 months good!
  if (nrow(df) == months){
    return(df)
  }
  # No data certain months. Fix it
  filled <- data.frame()
  for (m in 1:months){
    monthChar = as.character(m)
    if (m < 10){
      monthChar <- paste("0", monthChar, sep="")
    }
    newDf <- data.frame(month=monthChar, count=0)
    filled <- rbind(filled, newDf)
  }
  for (i in 1:nrow(df)){
    month = as.numeric(df$month[i])
    filled$count[month] <- df$count[i]
  }
  return(filled)
}

# Make a plot for each county
counties <- unique(result$county)
thisMonth = as.numeric(format(now_date, "%m"))
for(i in 1:length(counties)){
  countyN <- counties[i]
  countyPre <- preCovidYear %>% filter(county==countyN) %>% count(month) %>% rename(count=n)
  countyCovidYear <- covidYear %>% filter(county==countyN) %>% count(month) %>% rename(count=n)
  
  countyPre <- fill_empty_months(countyPre, 12)
  countyCovidYear <- fill_empty_months(countyCovidYear, thisMonth)

  plot <- ggplot() +
  geom_line(data=countyPre, aes(x=month, y=count, group=2019, col="2019")) +
  geom_line(data=countyCovidYear, aes(x=month, y=count, group=2020, col="2020")) +
  ggtitle(countyN) +
  ylab("Number of bankruptcies") + xlab("Month") 
  print(plot)
}
```

Based on the results no clear conclusion can be taken as there are no significant change. However looking at these numbers alone is not enough to give a final conclusion. This is because some company types are more affected than others such as the musical industry. 

It would thus be neccecary to include other data to be able to take a better conclusion.
