---
title: "BirthDataUS"
author: "Jorgen Reinnes"
date: "10/25/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library("plyr")
library("magrittr")
library("tidyverse")
library("ggplot2")
knitr::opts_chunk$set(echo = FALSE)
```

## Analysis of birth data from the us 2017 to 2019

```{r loadData}
readBirthData <- function(file, col_positions) {
  col_types <- cols(
    birthYear = col_number(),
    birthMonth = col_number(),
    birthDayOfWeek = col_number(),
    sex = col_character(),
    weight = col_number()
  )
  birthData <- read_fwf(
    file,
    col_positions,
    col_types,
    trim_ws = TRUE,
    skip = 0,
    n_max = Inf,
    # guess_max = min(n_max, 1000),
    progress = show_progress(),
    skip_empty_rows = TRUE
  )
}

# All data use the same positional data for columns
# 475 = Gender
# 13-14 = DOB_MM  (MM)
# 504-507 = birth weight in grams EXCLUDE where 9999 as that is unstated weight!
# 23 = birth day of week    (d)
col_positions <- fwf_positions(c(9, 13, 23, 475, 504), c(12, 14, 23, 475, 507), c("birthYear", "birthMonth", "birthDayOfWeek", "sex", "weight"))
result <- data.frame()
# 2018
result <- rbind(result, readBirthData("data/Nat2017PublicUS.c20180516.r20180808.txt", col_positions))
# 2018
result <- rbind(result, readBirthData("data/Nat2018PublicUS.c20190509.r20190717.txt", col_positions))
# 2019
result <- rbind(result, readBirthData("data/Nat2019PublicUS.c20200506.r20200915.txt", col_positions))
```

### Proportion of boys to girls

```{r proportionOfBoysToGirls}

# Aggregate number of boys and girls each month

# Calculate male/female proportion grouped by birthYear and month
proportionData <- result %>% group_by(birthYear, birthMonth) %>% count(sex) %>%
  # Calculate proportion ex maleCount / ((male + female count) - male count)
  mutate(proportion = n / (sum(n)-n)) %>%
  # Filter out all female
  filter(sex == "M") %>%
  # Re group by year
  ungroup() %>%
  group_by(birthYear)

# Plot 
  ggplot() +
  geom_line(data=proportionData, aes(x=birthMonth, y=proportion, group=birthYear, col=as.character(birthYear))) +
  ggtitle("Proportion of boys/girls births") +
  ylab("Proportion") + xlab("Month") +
  labs(color='Year') +
  scale_x_continuous(expand = c(0, 0))
  
  moreBornBoys <- floor((mean(proportionData$proportion)-1)*100)
  
```

Based on the given data there were born r moreBornBoys`% more boys than girls from 2017 to 2019. The change in proportion is going downwards towards December.





### The average birth weight in grams by gender?

```{r averageBirthWeightVersusGender}
# Calculate male/female avg birth weight across months
weightData <- result %>% group_by(birthYear, birthMonth, sex) %>%
  # Calculate proportion ex maleCount / ((male + female count) - male count)
  mutate(avgWeight = mean(weight)) %>%
  distinct(birthYear, birthMonth, sex, .keep_all = TRUE) %>%
  mutate(yearAndGender = paste(birthYear, sex, sep=": ")) %>%
  # Re group by year
  ungroup() %>%
  group_by(birthYear)

# Plot 
  ggplot() +
  geom_line(data=weightData, aes(x=birthMonth, y=avgWeight, linetype=sex, group=yearAndGender, col=as.character(birthYear))) +
  ggtitle("Average Birth Weight by Gender") +
  ylab("Weight (g)") + xlab("Month") +
  labs(color='Year', linetype="Sex") + 
    scale_x_continuous(expand = c(0, 0))

```

On average males weigh more than girls and the trend has been stable across the years. However of note weight is marginally higher in the summer than it is during the winter.



### The proportion of boys to girls by day of birth.


```{r proportionByDayOfWeek}
# Calculate male/female proportion by weekday per year
proportionDataWeekday <- result %>% group_by(birthDayOfWeek, birthYear) %>% count(sex) %>%
  # Calculate proportion ex maleCount / ((male + female count) - male count)
  mutate(proportion = n / (sum(n)-n)) %>%
  mutate(color = "#EFC000FF") %>%
  # Fix that americans use sunday as 0th day of the week
  mutate(birthDayOfWeek = case_when(birthDayOfWeek == 0 ~ 7, 
                             TRUE ~ birthDayOfWeek)) %>%
  # Filter out all female
  filter(sex == "M") %>%
  # Re group by year
  ungroup() %>%
  group_by(birthYear)
  
  ggplot(proportionDataWeekday, aes(x = birthDayOfWeek, y = proportion)) +
  geom_bar(
    aes(group=as.character(birthYear), fill=as.character(birthYear)),
    stat = "identity", position = position_dodge(0.8),
    width = 0.7
    ) +
    ylab("Proportion") + xlab("Day of the week with 1 == monday and sunday == 7") +
    labs(fill='Year')+ 
    scale_y_continuous(expand = c(0, 0))
```


The proportion between boys and girls born does not seem to depend a noticeable amount each yea on which weekday they are born on. 















